---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  out.width = "100%"
)

# library(noclocksai)
pkgload::load_all()

library(ellmer)
library(DiagrammeR)
```

```{r load_cached_resources, echo=FALSE, include=FALSE}
mermaid_vignette_resources <- qs2::qs_read(pkg_sys_extdata("vignettes/mermaid_vignette_resources.qs"))
mermaid_code <- mermaid_vignette_resources$mermaid_code
resp <- mermaid_vignette_resources$resp
mermaid_workflow_code <- mermaid_vignette_resources$mermaid_workflow_code
resp_workflow <- mermaid_vignette_resources$resp_workflow
```

## Example: Mermaid.js Diagram Generation

This example demonstrates how to use `noclocksai`'s features that build on top of the `ellmer`
package to generate a [Mermaid.js](https://mermaid-js.github.io/mermaid/#/) diagram based on a provided R code snippet. 
The `ellmer` package provides a simple interface to interact with the OpenAI API, allowing you to generate
the diagrams based on user-provided context or code snippets.

`noclocksai` comes with a set of pre-defined prompts for generating Mermaid.js diagrams, making it easy to
generate diagrams without having to write the prompts from scratch.

In this example the following steps will be taken:

1. Initialize a chat with OpenAI using the Mermaid.js system prompt.
2. Pass an example R code snippet to the chat.
3. Extract the generated Mermaid.js diagram code block from the response.
4. Render the Mermaid.js diagram using the extracted code block.

### Prompts

The `noclocksai` package provides two prompts for generating Mermaid.js diagrams. These prompts can be found in the package's `prompts/mermaid/` installed directory and retrieved via the `noclocksai::pkg_sys_prompt()` utility function or directly called using the `prompt_mermaid_sys()` and `prompt_mermaid_user()` functions.

Expand the following sections to view the Mermaid.js system and user prompts, respectively.

<br>

<details><summary>View [Mermaid.js System Prompt](../inst/prompts/mermaid/system.prompt.md)</summary><p>

````markdown

```{r child="../inst/prompts/mermaid/system.prompt.md"}
```

````

</p></details>

<br>

<details><summary>View [Mermaid.js User Prompt](../inst/prompts/mermaid/user.prompt.md)</summary><p>

````markdown

```{r child="../inst/prompts/mermaid/user.prompt.md"}
```

````

</p></details>

### Diagram Generation Workflow

The following code demonstrates how to generate a Mermaid.js diagram based on a provided R code snippet
using the `noclocksai` package:

```{r mermaid_diagram_generation_chat, eval=FALSE}
# initialize a chat with OpenAI using the mermaid.js system prompt
chat <- ellmer::chat_openai(
  system_prompt = prompt_mermaid_sys(),
  echo = TRUE,
  api_args = list(temperature = 0)
)

# example code to pass to chat
example <- "starwars |>
  group_by(species) |>
  summarise(
    n = n(),
    mass = mean(mass, na.rm = TRUE)
  ) |>
  filter(
    n > 1,
    mass > 50
  )"

# send the example code to the chat
resp <- chat$chat(prompt_mermaid_user(context = example))
```

The response from the chat will contain the generated Mermaid.js diagram code block:

```{r mermaid_diagram_generation_chat_output, echo=FALSE}
cat(resp, sep = "\n")
```

Now that the chat has been initialized and the example R code snippet has been passed to the chat,
we can extract the generated Mermaid.js diagram code block from the response using the
provided utility function `extract_code()`:

```{r extract_mermaid_diagram_code, eval=FALSE}
# extract mermaid diagram code block from response
mermaid_code <- extract_code(resp, lang = "mermaid", print = TRUE)
```

preview the extracted Mermaid.js diagram code block:

```{r extract_mermaid_diagram_code_output, echo=FALSE}
cat(mermaid_code, sep = "\n")
```

Then you can use this generated Mermaid.js code block to render a diagram for the example R code:

```{r render_mermaid_diagram, out.height="400px", fig.dim=c(6,4), fig.align="center", fig.cap="Mermaid.js Diagram for the Example R Code Snippet"}
# render the mermaid diagram
DiagrammeR::DiagrammeR(mermaid_code)
```

<br>

The generated Mermaid.js diagram should represent the flow of the provided R code snippet.

### Mermaid for this Workflow

Let's generate a diagram for the workflow defined above:

```{r mermaid_workflow, eval=FALSE}
workflow <- "
1. Initialize a chat with OpenAI using the mermaid.js system prompt
2. Pass an example R code snippet to the chat
3. Extract the generated Mermaid.js diagram code block from the response
4. Render the Mermaid.js diagram using the extracted code block
"

resp <- chat$chat(prompt_mermaid_user(context = workflow))
```

The response from the chat will contain the generated Mermaid.js diagram code block:

```{r mermaid_workflow_output, eval=FALSE}
cat(resp_workflow, sep = "\n")
```

Extract the generated Mermaid.js diagram code block:

```{r extract_mermaid_workflow_code, eval=FALSE}
# extract mermaid diagram code block from response
mermaid_workflow_code <- extract_code(resp, lang = "mermaid", print = TRUE)
```

preview the extracted Mermaid.js diagram code block:

```{r extract_mermaid_workflow_code_output, eval=FALSE}
cat(mermaid_workflow_code, sep = "\n")
```

Render the Mermaid.js diagram for this workflow:

```{r render_mermaid_workflow, out.height="400px", fig.dim=c(6,4), fig.align="center", fig.cap="Mermaid.js Diagram for the Workflow of Generating a Mermaid.js Diagram"}
# render the mermaid diagram
DiagrammeR::DiagrammeR(mermaid_workflow_code)
```

<br>

The generated Mermaid.js diagram should represent the workflow of generating a Mermaid.js diagram based on an R code snippet.
